---
title: "Workflow"
author: "Aleix"
date: "Wednesday, July 02, 2014"
output: pdf_document
---

# Functions

## General:

* Be explicit and descriptive with names, other people will read your code!
* Syntax: `myFunction_task` or `my.function_task`. The `_` separates the name from the task.
* `optL <- list()` is a list where you put all the parameters that you need. 
When put in the header `myFunction(..., optL = list(), ...)` mean that it is optional for the call.
* Each file should start with a small description of what it does


## Data transformations
* Function header: `myFunction_transform <- function(bikeData, optL = list())`
* Folder: 'transform'
* File name: 'myFunction_transform.R'
* Value: data.frame

## Modeling functions
* Function header: `myFunction_model <- function(historicBikeData, optL = list())`
* Function header: `myFunction_predict <- function(toPredictionBikeData, historicModel)`
* Folder: 'model'
* File name: 'myFunction_model.R' (both functions in the same file)
* Value of myFunction_model: can be everything, including a list of objects
* Value of myFunction_predict: vector, matrix or data.frame


## Cross validation functions and others
* Function header: `myFunction_cv <- function(do_model, do_prediction, do_transform, bikeData, optL = list(kFold = 5))`
* Folder: 'general'
* File name: 'myFunction_cv.R' (both functions in the same file)
* Value: a numeric vector


# Example

_Transformation_ : We complete the data set with expliciting time information
```{r}
completeTime_transform <- function(bikeData){
  bikeData$datetime <- ymd_hms(bikeData$datetime)
  bikeData$day <- as.Date(bikeData$datetime)
  bikeData$hour <- hour(bikeData$datetime)
  bikeData$month <- month(bikeData$datetime)
  bikeData$weekDay <- wday(bikeData$datetime)
  
  return(bikeData)
}
```

_Model_ : This model splits by categories in variables and computes the mean

```{r}
categoryMeans_model <- function(historicBikeData, optL = list()){
  categoryModel <- ddply(historicBikeData, .(season, hour, workingday), function(splitDataSet){
    mean(splitDataSet$count)    
  })
  colnames(categoryModel)[4] <- "mean"
  
  return(categoryModel)
}

categoryMeans_predict <- function(toPredictionBikeData, historicModel){
  apply(toPredictionBikeData, 1, function(rowData){
    obsSeason <- as.numeric(rowData['season'])
    obsHour <- as.numeric(rowData['hour'])
    obsWorkingDay <- as.numeric(rowData['workingday'])
    selecIndex <- historicModel$season == obsSeason & historicModel$hour == obsHour & historicModel$workingday == obsWorkingDay
    foundCategory <- ( sum(selecIndex) > 0 )
    if( foundCategory ){
      predictionValue <- historicModel$mean[selecIndex]
    }else{
      predictionValue <- NA
    }
  })
}
```

_Loss function_ : Kaggle's competition loss function

```{r}
kaggle_loss <- function(realValues, predictedValues){
  sqrt( sum( na.omit( ( log( realValues + 1 ) - log( predictedValues + 1 ) )^2 ) ) )
}
```


_Cross validation_ : This cross validation splits the data set randomly

```{r}
random_cv <- function(do_model, do_prediction, do_transform, bikeData, optL = list(kFold = 5)){
  
  lengthSplit <- floor( nrow(bikeData)/optL$kFold )
  
  sapply(1:optL$kFold, function(cvIter){
    trainIndex <- sample(1:nrow(bikeData), size = lengthSplit, replace = FALSE)
    trainSet <- bikeData[ trainIndex, ]
    testSet <- bikeData[ -trainIndex, ]
    
    trainSet <- do_transform( trainSet ) 
    ## this is important to be done here instead out of the loop to ensure the good working of the cross valdation (not using information of the test set)
    newModel <- do_model( trainSet, optL )
    
    testSet <- do_transform( testSet )
    newPredictions <- do_prediction( testSet, newModel )
    realValues <- testSet$count
    
    return( kaggle_loss( realValues, newPredictions ) )
  })
}
```


_Execution_
```{r}
require(lubridate)
require(plyr)
bikeData <- read.csv('c:/Users/Aleix/Google Drive/Archivos/KaggleBike/train.csv', head = TRUE, stringsAsFactors = FALSE)

random_cv(categoryMeans_model, categoryMeans_predict, completeTime_transform, bikeData)
```